---
title: "Social Networks - Comparing Dyad Census in Simulated Networks"
author: "Vullings, Niels"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    theme: journal
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clear cache

```{r}
rm(list=ls())
```

### Check packages

```{r}
#
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}
```

## Load packages

```{r}
packages <- c("igraph", "RsienaTwoStep")
fpackage.check(packages)

# assign(paste0("blabla", date()), t1)
```

## Load simulation Data

```{r}
# getwd()

load("articles/data/processed/20231108ts_ansSp3.rda") #SimStep
ts_ansSp3 <- x
rm(x)

load("articles/data/processed/20231114ts_ansSCp3.rda") #Strict Coordination
ts_ansSCp3 <- x
rm(x)

load("articles/data/processed/20231112ts_ansWCp3.rda") #Strict Coordination
ts_ansWCp3 <- x
rm(x)  

```

# Old Functions

## Dyad census function for observed network evolution

```{r}
dyad.census.ind <- function(T1, T2, save.table = FALSE, table.name = "simsamsom") {
  dft1 <- reshape2::melt((T1 + t(T1)))[as.vector(upper.tri(T1)),]
  dft1 <- dft1[order(dft1$Var1),]
  rownames(dft1) <- NULL
  colnames(dft1) <- c("ego", "alter", "tie")

  dft2 <- reshape2::melt((T2 + t(T2)))[as.vector(upper.tri(T2)),]
  dft2 <- dft2[order(dft2$Var1),]
  rownames(dft2) <- NULL
  colnames(dft2) <- c("ego", "alter", "tie")

  dft1$tiet2 <- dft2$tie

  tR <- table(dft1$tie, dft1$tiet2)
  colnames(tR) = c("Null", "Assym", "Mutual")
  rownames(tR) = c("Null", "Assym", "Mutual")


  if(save.table == TRUE){


    print("Dyad transitions from T1 to T2")
    assign(paste0(table.name), dft1, envir = .GlobalEnv)
    # assign(paste0(table.name, ": " , format(as.POSIXct(Sys.time()), format = "%H:%M:%S")), dft1, envir = .GlobalEnv)

    return(tR)

  } else{

    ##Give Output
    print("Dyad transitions from T1 to T2")
    return(tR)
  }

}

```

## Dyad census function for simulated data using simnets

```{r}

# dyad.census.sim.ind <- function(T1, T_sim, sim_sample = c(1:100), simN = 5, seed = 345657343, save.table = FALSE, table.name = "simsamsom"){
#   ##t1 related census
#   dft1 <- reshape2::melt((T1 + t(T1)))[as.vector(upper.tri(T1)),]
#   dft1 <- dft1[order(dft1$Var1),]
#   rownames(dft1) <- NULL
#   colnames(dft1) <- c("ego", "alter", "tie")
#   
#   ##t2 related census with sim
#   
#   ###Set seed and create an random sample for simulations
#   set.seed(seed)
#   if(simN > length(sim_sample)){ 
#     print("Warning, simN higher than the number of simulations")
#   } else{
#     
#     sample <- as.vector(sample(sim_sample, simN, replace = TRUE))
#   }
#   
#   sim_tot <- 0 #start sim_tot with 0 to solve NA error
#   
#   for (x in sample) {
#     
#     sim_temp <- reshape2::melt((T_sim$simnets[[x]] + t(T_sim$simnets[[x]])))[as.vector(upper.tri(T_sim$simnets[[x]])),]
#     sim_tot <- sim_tot + sim_temp 
#     
#     x+1
#     
#   }
#   sim_tot <<- sim_tot 
#   sim_mean <<- sim_tot/length(sample)
#   sim_mean <- round(sim_mean)
#   sim_mean <- sim_mean[order(sim_mean$Var1),]
#   rownames(sim_mean) <- NULL
#   colnames(sim_mean) <- c("ego", "alter", "tie")
#   
#   
#   #Create table
#   dft1$tiet2 <- sim_mean$tie
#   tR <- table(dft1$tie, dft1$tiet2)
#   
#   colnames(tR) = c("Null", "Assym", "Mutual")
#   rownames(tR) = c("Null", "Assym", "Mutual")
#   
#   #give the option to save the table in global environment
#   if(save.table == TRUE){
#     
#     # Table <<-  dft1
#     print("Dyad transitions from T1 to simulated T2")    
#     assign(paste0(table.name), dft1, envir = .GlobalEnv)
#     # assign(paste0(table.name, ": " , format(as.POSIXct(Sys.time()), format = "%H:%M:%S")), dft1, envir = .GlobalEnv)
#     
#     return(tR)
#     
#   } else{
#     
#     print("Dyad transitions from T1 to simulated T2")
#     return(tR)
#     
#     
#   }
#   
# }
# ?Sys.time
```

## Dyad census function for simulated data using simnets [different method]

```{r}
# 
# dyad.census.sim.ind2 <- function(T1, T_sim, sim_sample = c(1:100), simN = 5, seed = 345657343, save.table = FALSE, table.name = "simsamsom"){
#   ##t1 related census
#   dft1 <- reshape2::melt((T1 + t(T1)))[as.vector(upper.tri(T1)),]
#   dft1 <- dft1[order(dft1$Var1),]
#   rownames(dft1) <- NULL
#   colnames(dft1) <- c("ego", "alter", "tie")
#   
#   ##t2 related census with sim
#   
#   ###Set seed and create an random sample for simulations
#   set.seed(seed)
#   if(simN > length(sim_sample)){ 
#     print("Warning, simN higher than the number of simulations")
#   } else{
#     
#     sample <- as.vector(sample(sim_sample, simN, replace = TRUE))
#   }
#   
#   sim_tot <- 0 #start sim_tot with 0 to solve NA error
#   
#   for (x in sample) {
#     
#     sim_temp <- T_sim$simnets[[x]]
#     sim_tot <- sim_tot + sim_temp 
#     
#     x+1
#     
#   }
#   
#   sim_tot <<- sim_tot 
#   sim_mean <<- sim_tot/length(sample)
#   T2 <<- round(sim_mean)
#   dft2 <<- reshape2::melt((T2 + t(T2)))[as.vector(upper.tri(T2)),]
#   
#   dft2 <- dft2[order(dft2$Var1),]
#   rownames(dft2) <- NULL
#   colnames(dft2) <- c("ego", "alter", "tie")
#   #
#   #
#   # #Create table
#   dft1$tiet2 <- dft2$tie
#   tR <<- table(dft1$tie, dft1$tiet2)
#   
#   colnames(tR) = c("Null", "Assym", "Mutual")
#   rownames(tR) = c("Null", "Assym", "Mutual")
#   
#   #give the option to save the table in global environment
#   if(save.table == TRUE){
#     
#     # Table <<-  dft1
#     print("Dyad transitions from T1 to simulated T2")
#     assign(paste0(table.name), dft1, envir = .GlobalEnv)
#     # assign(paste0(table.name, ": " , format(as.POSIXct(Sys.time()), format = "%H:%M:%S")), dft1, envir = .GlobalEnv)
#     
#     return(tR)
#     
#   } else{
#     
#     print("Dyad transitions from T1 to simulated T2")
#     return(tR)
#     
#     
#   }
#   
# }
# ?Sys.time
```

# Improved functions The functions below are improved and are able to
measure specific dyad types in a social network.

## Dyad Census type function [improved method]

The function dyad.census.type allows for exporting a dataframe that
lists the specific a dyad's relationship evolved between the 2 point of
observation. Its standard function is to count the occurences of certain
dyad relationships.

```{r}

dyad.census.type <- function(t1, t2, save.dyads = FALSE){
    
    t1 <- t1 #+ 1
    diag(t1) = NA #exclude the diagonal, not relevant data
    t2 <- t2
    diag(t2) = NA #exclude the diagonal, not relevant data
    
    flips <- t1 + t(t2) - t2 #subtract s502 to ensure that Mutual ties are not included
    
    
    jumpst1 <- t1 + t(t1) #t1 plus its transpose lead to a value of 2 for mutual ties
    stablet1 <- jumpst1 #use this for stable assymetric ties
    jumpst1[lower.tri(jumpst1)] <- NA #remove duplicate ties
    
    jumpst2 <- t2 + t(t2)
    stablet2 <- jumpst2 #use this for stable assymetric ties
    jumpst2[lower.tri(jumpst2)] <- NA #remove duplicate ties
    
    #Dyad combinations
    stable00 <<- as.data.frame(which(jumpst1 == 0 & jumpst2 == 0, arr.ind = TRUE)) #Null at t1 and Null at t2
    stable01 <<- as.data.frame(which(flips == 0 & stablet1 == 1 & stablet2 == 1, arr.ind = TRUE)) #Assymetric at t1 and t2 and no flip
    stable11 <<- as.data.frame(which(jumpst1 == 2 & jumpst2 == 2, arr.ind = TRUE)) #Mutual at t1 and Mutual at t2
    
    Null_Assym <<- as.data.frame(which(jumpst1 == 0 & stablet2 == 1, arr.ind = TRUE))#Null -> assym
    Assym_Null <<- as.data.frame(which(flips == 0 & stablet1 == 1 & stablet2 == 0, arr.ind = TRUE))#Assum -> Null
    Assym_Mut <<- as.data.frame(which(stablet1 == 1 & jumpst2 == 2, arr.ind = TRUE))#Assym -> Mutual
    Mut_Assym <<- as.data.frame(which(jumpst1 == 2 & jumpst2 == 1, arr.ind = TRUE))#Assym -> Mutual
    
    flip <<- as.data.frame(which(flips == 2 & stablet1 != 2 & stablet2 != 2, arr.ind = TRUE)) #flips, so 01 at T1 and 10 at T2
    jump02 <<- as.data.frame(which(jumpst1 == 0 & jumpst2 == 2, arr.ind = TRUE)) #Null jump, from Null to Mutual
    jump20 <<- as.data.frame(which(jumpst1 == 2 & jumpst2 == 0, arr.ind = TRUE)) #Mutual jump, from Mutual to Null
    
    
    table <- cbind(nrow(stable00),
                   nrow(stable01),
                   nrow(stable11),
                   nrow(Null_Assym),
                   nrow(Assym_Null),
                   nrow(Assym_Mut),
                   nrow(Mut_Assym),
                   nrow(flip),
                   nrow(jump02),
                   nrow(jump20)) #bind the rownumbers to determine the census of the tie type
    # print(table)
    colnames(table) = c("Null > Null", 
                        "Assym > Assym",
                        "Mutual > Mutual",
                        "Null > Assym",
                        "Null > Assym",
                        "Assym > Mutual",
                        "Mutual > Assym",
                        "Tie flip",
                        "Null > Mutual",
                        "Mutual > Null") 
    
    if(save.dyads == TRUE){
      #give each dyad a relationship identifier
      stable00$tie <- ifelse(is.na(stable00$row),NA, "Null > Null")
      stable01$tie <- ifelse(is.na(stable01$row),NA, "Assym > Assym")
      stable11$tie <- ifelse(is.na(stable11$row),NA, "Mutual > Mutual")
      Null_Assym$tie <- ifelse(is.na(Null_Assym$row),NA, "Null > Assym")
      Assym_Null$tie <- ifelse(is.na(Assym_Null$row),NA, "Assym > Null")
      Assym_Mut$tie <- ifelse(is.na(Assym_Mut$row),NA, "Assym > Mutual")
      Mut_Assym$tie <- ifelse(is.na(Mut_Assym$row),NA, "Mutual > Assym")
      flip$tie <- ifelse(is.na(flip$row),NA, "Tie flip")
      jump02$tie <- ifelse(is.na(jump02$row),NA, "Null > Mutual")
      jump20$tie <- ifelse(is.na(jump20$row),NA, "Mutual > Null")
      
      dyads <<- rbind(stable00,stable01,stable11, Null_Assym, Assym_Null, Assym_Mut, Mut_Assym, flip,jump02,jump20) #save a dataframe which contains the relationships of each dyad
      
      # print("Dyad transitions from T1 to T2")
      return(as.data.frame(table))
      
    } else{
      
      # print("Dyad transitions from T1 to T2")
      return(as.data.frame(table))
      
      
    }

}


# sna::dyad.census(s501,s502)

# dyad.census.type(s501[1:5,1:5],s502[1:5,1:5], save.dyads = T)

```

## Dyad simulated census function

This function returns the dyad census for simulated data. Simulated
networks can be included as t2. The function will create a mean census
for the simulate data, based on specified parameters.

```{r}

dyad.census.sim <- function(t1, tsim, sim_sample = c(1:100), simN = 5, seed = 345657343){

  set.seed(seed)
  if(simN > length(sim_sample)){
    print("Warning, simN higher than the number of simulations")
  } else{

    sample <- as.vector(sample(sim_sample, simN, replace = TRUE))
  }


  df <- foreach::foreach(sim_sample, i=iterators::icount(), .combine="rbind") %dopar% {
    df <- as.data.frame(dyad.census.type(s501, t2= ts_ansSCp3$simnets[[i]]))
  }

  assign("testje", df, envir = .GlobalEnv)
  test <- round(colMeans(df))
  return(test)
}

# dyad.census.sim(s501,ts_ansSCp3$simnets)
```

## Dyad census full function for observed and simulated data

This is the complete function that can handle both observed and
simulated data. it integrates the functions that were described above,
so that single matrices as well as lists (sim data) can be included in
the census.

things to fix:

\- If a dyad type is not present, the code will crash because there is
an empty dataframe [Fixed]
\- Return output in Matrix format for readibility

```{r}
ts_dyad.census <- function(t1,t2, sim_sample = c(1:1000), simN = 1000, seed = 345657343, save.dyads = FALSE){
  
  save <- save.dyads #savethe boolean for later
  #inner function for dyad census 
  dyad.census.type <- function(t1, t2, save.dyads = FALSE){
    
    t1 <- t1 #+ 1
    diag(t1) = NA #exclude the diagonal, not relevant data
    t2 <- t2
    diag(t2) = NA #exclude the diagonal, not relevant data
    
    flips <- t1 + t(t2) - t2 #subtract s502 to ensure that Mutual ties are not included
    
    
    jumpst1 <- t1 + t(t1) #t1 plus its transpose lead to a value of 2 for mutual ties
    stablet1 <- jumpst1 #use this for stable assymetric ties
    jumpst1[lower.tri(jumpst1)] <- NA #remove duplicate ties
    
    jumpst2 <- t2 + t(t2)
    stablet2 <- jumpst2 #use this for stable assymetric ties
    jumpst2[lower.tri(jumpst2)] <- NA #remove duplicate ties
    
    #Dyad combinations
    stable00 <- as.data.frame(which(jumpst1 == 0 & jumpst2 == 0, arr.ind = TRUE)) #Null at t1 and Null at t2
    stable01 <- as.data.frame(which(flips == 0 & stablet1 == 1 & stablet2 == 1, arr.ind = TRUE)) #Assymetric at t1 and t2 and no flip
    stable11 <- as.data.frame(which(jumpst1 == 2 & jumpst2 == 2, arr.ind = TRUE)) #Mutual at t1 and Mutual at t2
    
    Null_Assym <- as.data.frame(which(jumpst1 == 0 & stablet2 == 1, arr.ind = TRUE))#Null -> assym
    Assym_Null <- as.data.frame(which(flips == 0 & stablet1 == 1 & stablet2 == 0, arr.ind = TRUE))#Assum -> Null
    Assym_Mut <- as.data.frame(which(stablet1 == 1 & jumpst2 == 2, arr.ind = TRUE))#Assym -> Mutual
    Mut_Assym <- as.data.frame(which(jumpst1 == 2 & jumpst2 == 1, arr.ind = TRUE))#Assym -> Mutual
    
    flip <- as.data.frame(which(flips == 2 & stablet1 != 2 & stablet2 != 2, arr.ind = TRUE)) #flips, so 01 at T1 and 10 at T2
    jump02 <- as.data.frame(which(jumpst1 == 0 & jumpst2 == 2, arr.ind = TRUE)) #Null jump, from Null to Mutual
    jump20 <- as.data.frame(which(jumpst1 == 2 & jumpst2 == 0, arr.ind = TRUE)) #Mutual jump, from Mutual to Null
    
    
    table <- cbind(nrow(stable00),
                   nrow(stable01),
                   nrow(stable11),
                   nrow(Null_Assym),
                   nrow(Assym_Null),
                   nrow(Assym_Mut),
                   nrow(Mut_Assym),
                   nrow(flip),
                   nrow(jump02),
                   nrow(jump20)) #bind the rownumbers to determine the census of the tie type
    
    colnames(table) = c("Null > Null", 
                        "Assym > Assym",
                        "Mutual > Mutual",
                        "Null > Assym",
                        "Null > Assym",
                        "Assym > Mutual",
                        "Mutual > Assym",
                        "Tie flip",
                        "Null > Mutual",
                        "Mutual > Null") 
    
    if(save.dyads == TRUE){
      #give each dyad a relationship identifier
      stable00$tie <- ifelse(is.na(stable00$row),NA, "Null > Null")
      stable01$tie <- ifelse(is.na(stable01$row),NA, "Assym > Assym")
      stable11$tie <- ifelse(is.na(stable11$row),NA, "Mutual > Mutual")
      Null_Assym$tie <- ifelse(is.na(Null_Assym$row),NA, "Null > Assym")
      Assym_Null$tie <- ifelse(is.na(Assym_Null$row),NA, "Assym > Null")
      Assym_Mut$tie <- ifelse(is.na(Assym_Mut$row),NA, "Assym > Mutual")
      Mut_Assym$tie <- ifelse(is.na(Mut_Assym$row),NA, "Mutual > Assym")
      flip$tie <- ifelse(is.na(flip$row),NA, "Tie flip")
      jump02$tie <- ifelse(is.na(jump02$row),NA, "Null > Mutual")
      jump20$tie <- ifelse(is.na(jump20$row),NA, "Mutual > Null")
      
      dyads <<- rbind(stable00,stable01,stable11, Null_Assym, Assym_Null, Assym_Mut, Mut_Assym, flip,jump02,jump20) #save a dataframe which contains the relationships of each dyad

      # print("Dyad transitions from T1 to T2")
      return(as.data.frame(table))
      
    } else{
      
      # print("Dyad transitions from T1 to T2")
      return(as.data.frame(table))
      
      
    }
  }
  
  if(is.list(t2)){
    
    set.seed(seed) #set seed for similar outcomes with the simulations
    if(simN > length(sim_sample)){ 
      print("Warning, simN higher than the number of simulations")
    } else{
      
      sample <- as.vector(sample(sim_sample, simN, replace = TRUE)) #create a vector that chooses simulations at random
    }
    
    
    df <- foreach::foreach(sim_sample, i=iterators::icount(), .combine="rbind") %dopar% {
      df <- as.data.frame(dyad.census.type(t1, t2= t2[[i]])) #loop the dyad.census.type function over all relevant simulations and save in df
    }
    
    assign("Simulated ties", df, envir = .GlobalEnv) #save the census df to the global environment
    output <- round(colMeans(df)) #Take the mean of the dyad occurrences to get the census mean
    print(class(output))
    return(output)
  } else{
    dyad.census.type(t1,t2, save.dyads = save)
  }
}

# ts_dyad.census(s501,ts_ansSCp3$simnets, sim_sample = c(1:1000), simN = 1000)
# ts_dyad.census(s501,s502, save.dyads = T)
# ts_dyad.census(s501[1:3,1:3],s502[1:3,1:3], save.dyads = T)

# rowSums(`Simulated ties`)
# dyad.census.ind(s501,s502)

```


```{r}

```

